\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-09-25 gwnnDraft02.tex
edits (Dylan)

Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished in multifield inflation models~\cite{} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

with $\xi$ being the temperature ratio defined in Eq.~(\ref{eq:xi}). The bounds from effective number of neutrinos ~\citep{} mean that $\xi_i \lesssim 1$ for all $i$, so $g_{*} \approx g_{*,\gamma}$. In the case of dark QCD-like chiral  phase transitions, the temperature of the phase transition is on the the order of the symmetry breaking scale of the respective sector, $T_{h}^{i} \sim \mathcal{O}(\Lambda_{QCD,i})$. The work of~\citep{Bai:2018dxf} calculated $\alpha$ with an effective chiral Lagrangian have found various upper bounds. We take the optimistic scenario where the numerator is bounded above by the symmetry breaking scale

@@

Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). In this case, the efficiency factors for sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$. Quantities that are calculated at the time of nucleation are denoted with an asterisk (i.e $h^2\Omega_{\textrm{GW}}^*$), which are evolved to relate to their values at the time of observation. The form of the GW energy density at the time of nucleation is given by~\citep{Breitbach:2018ddu},

@@

\draftnote{I think we need more details on the scenarios that go into making those plots.} Assuming all else equal, we use the temperature schemes from Eq.~(\ref{eqn:energyDensityAllowed}) to produce the plots in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable. 

>>>>

Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished with primodial axion-like particle (ALP) models~\cite{MARSH20161,PhysRevLett.121.201303} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

with $\xi$ being the temperature ratio defined in Eq.~(\ref{eq:xi}). The bounds from effective number of neutrinos~\citep{Aghanim:2018eyx} mean that $\xi_i \lesssim 1$ for all $i$, so $g_{*} \approx g_{*,\gamma}$. In the case of dark QCD-like chiral  phase transitions, the temperature of the phase transition is on the the order of the symmetry breaking scale of the respective sector, $T_{h}^{i} \sim \mathcal{O}(\Lambda_{QCD,i})$. The work of~\citep{Bai:2018dxf} calculated $\alpha$ with an effective chiral Lagrangian have found various upper bounds. We take the optimistic scenario where the numerator is bounded above by the symmetry breaking scale

@@

Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$. Quantities that are calculated at the time of nucleation are denoted with an asterisk (i.e $h^2\Omega_{\textrm{GW}}^*$), which are evolved to relate to their values at the time of observation. The form of the GW energy density at the time of nucleation is given by~\citep{Breitbach:2018ddu},

@@

\draftnote{I think we need more details on the scenarios that go into making those plots.} Assuming all else equal, we use the temperature schemes from Eq.~(\ref{eqn:energyDensityAllowed}) to produce the plots in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. When relaxing the assumptions on $\beta/H$, the scenarios in figure~(\ref{fig:Haa}) are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$ The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable. 
